Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-04 conclusion.tex
conclusions. refinements.

More in general, this paper highlighted the fundamental role of WMS for
experimental science. HEP was amongst the first, if not the very first
experimental community to realize the importance of using WMS to manage their
computational campaign(s). As computing become increasingly critical for a range
of experiments, the impact of WMS on HEP foreshadows the importance of WMS for
other experiments (such as SKA, LSST etc.). % and their computational campaigns.
These experiments will have their own workload characteristics, resources types
and federation constraints, as well metrics of performance.  The design and
experience captured in this paper should prove invaluable for designing
computational campaigns and will provide a baseline to evaluate the relative
merits of different approaches.

The 52M core hours used by ATLAS, via PanDA, is over 2\% of the total
utilization on Titan over the same period, bringing the time-averaged
utilization of Titan to be consistently upwards of 90\%. Given that the
average average utilization of most other leadership class machines is less
(e.g., NSF's flagship Blue Waters the average utilization fluctuates between
60-80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
approaches elsewhere. These unprecedented efficiency gains aside, this work is
just a starting point towards more effective operational models for future
leadership and online analytical platforms~\cite{foap-url}. These platforms
will have to support ever increasing complex workloads with varying models for
dynamic resource federation. Recent efforts with Google Cloud for the CMS
experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
different approaches to dynamic resource federation.

Our experience demonstrated the advantages arising from a  systems engineering
and abstractions based approach, in particular the pilot abstraction. As the
type, heterogeneity and complexity of future platforms  (e.g., Exascale,
Commercial Clouds) for science increases, there will be a greater need and
emphasis on abstractions to ensure WMS can both utilize as well as support
platforms.

>>>>

More generally, this paper highlights the fundamental role of WMS for
experimental science. HEP was amongst the first, if not the very first
experimental community to realize the importance of using WMS to manage their
computational campaign(s). As computing becomes increasingly critical for a
range of experiments, the impact of WMS on HEP foreshadows the importance of
WMS for other experiments (such as SKA, LSST etc.). % and their computational
campaigns. These experiments will have their own workload characteristics,
resources types and federation constraints, as well metrics of performance.
The design and experience captured in this paper will prove invaluable for
designing WMS to support computational campaigns and will provide a baseline
to evaluate the relative merits of different approaches.

The 52M core hours used by ATLAS, via PanDA, is over 2\% of the total
utilization on Titan over the same period, bringing the time-averaged
utilization of Titan to be consistently upwards of 90\%. Given that the
average average utilization of most other leadership class machines is less
(e.g., NSF's flagship Blue Waters the average utilization fluctuates between
60-80\% (see XDMoD\cite{bw-sucks})) there is ample headroom for similar
approaches elsewhere. These unprecedented efficiency gains aside, this work is
just a starting point towards more effective operational models for future
leadership and online analytical platforms~\cite{foap-url}. These platforms
will have to support ever increasing complex workloads with varying models for
dynamic resource federation. Recent efforts with Google Cloud for the CMS
experiment -- HEPCloud~\cite{hepcloud,googlehep} represent similar yet
different approaches to dynamic resource federation. As the type,
heterogeneity and complexity of future platforms  (e.g., Exascale, Commercial
Clouds) for science increases, there will be a greater need and emphasis on
abstractions to ensure WMS can both utilize as well as support evolving
platforms. Our experience demonstrated the advantages arising from an
abstractions based approach \textendash{} in particular scalable implementations
of pilot abstraction.
